---
title: "1.Acquisition of bioactive compounds, targets and PCT Network Construction"
format: 
  wordcount-html:
    toc: true
    number-sections: true
---

# Acquisition of bioactive compounds

## Load packages {-}
```{r}
library(tidyverse)
library(data.table)
library(openxlsx)

library(future)     
library(furrr)      
library(progressr) 
library(httr)
```

## Blood-absorbed bioactive compounds

Load Raw Compound Data from DCABM-TCM Database

```{r}
rm(list = ls())

source("utils.R")

compounds_dcabm <- read.xlsx("../data/raw/qfd/compounds_dcabm.xlsx") %>%
  dplyr::select(
    `Medicinal.materials(Chinese.Pinyin.name)`,
    `Medicinal.materials(Latin.name)`, 
    Constituents) %>%
  mutate(Compounds = Constituents) %>%
  # Trim leading/trailing whitespace from all character columns
  mutate(across(where(is.character), trimws)) %>%
  
  # Split multiple compounds separated by "|" into individual rows
  separate_rows(Compounds, sep = "\\|") %>%
  
  # Extract core metadata from the unstructured Compounds string
  mutate(
    CID = str_extract(Compounds, "(?<=Pubchem CID:)[^)]+") %>% str_trim(),
    DCABM.ID = str_extract(Compounds, "(?<=DCABM ID:)[^)]+") %>% str_trim(),
    temp = str_remove(Compounds, "\\s+\\(Pubchem CID:[^)]+\\)")
  ) %>%
  
  # Extract clean compound name and molecular formula
  mutate(
    temp2 = str_remove(temp, "\\([^)]+\\)$") %>% str_trim(),
    Compound = str_remove(temp2, "\\([^)]+\\)$") %>% str_trim(),
    Formula = str_match(temp2, "\\(([^)]+)\\)$")[, 2] %>% str_trim()
  ) %>%
  
  # Remove redundant intermediate columns
  dplyr::select(-Compounds, -temp, -temp2, -DCABM.ID) %>%
  
  # Final whitespace standardization for all character columns
  mutate(across(where(is.character), trimws)) %>%
  dplyr::select(-Constituents) %>%
  mutate(Source = "DCABM-TCM") %>%
  dplyr::select(
    `Medicinal.materials(Latin.name)`,
    `Medicinal.materials(Chinese.Pinyin.name)`,
    Compound,
    Formula,
    CID,
    Source
  )
``` 

## Load Raw Compound Data from literature
```{r}
compounds_qfd <-  read.xlsx("../data/raw/qfd/compounds_literature.xlsx")
```

## All compounds
```{r}
compounds_all <- compounds_dcabm %>%
  dplyr::select(`Medicinal.materials(Latin.name)`,
                `Medicinal.materials(Chinese.Pinyin.name)`, 
                Compound, 
                Formula, 
                CID, 
                Source
  ) %>%
  rbind(compounds_qfd) %>%
  group_by(Compound) %>%
  mutate(
    Compound_Count = n(),                         
    Source.new = paste(unique(Source), collapse = "; ") 
  ) %>%
  ungroup() %>%
  dplyr::select(-Source) %>%
  dplyr::rename(Source = Source.new) %>%
  distinct(Compound, .keep_all = TRUE)

# Save cleaned dataset
save(
  compounds_all,
  file = "../data/processed/01_compounds_all.RData"
)

openxlsx::write.xlsx(compounds_all,"../results/01_compounds_all.xlsx")
```

## Process
```{r}
compounds_all_processed <- openxlsx::read.xlsx("../results/01_compounds_all_manu_process.xlsx")
```

## Batch Retrieve PubChem Compound Annotations
```{r}
compounds_pubchem <- compounds_all_processed %>%
  mutate(`CID` = clean_cids(CID)) %>%
  filter(!is.na(CID)) %>%
  distinct(CID, .keep_all = TRUE)

properties <- c("IUPACName", "MolecularFormula", "MolecularWeight",
                "SMILES")

# Split unique CIDs into batches of 20 to comply with PubChem API rate limits
unique_cids <- unique(compounds_pubchem$CID)
cid_batches <- split(unique_cids, ceiling(seq_along(unique_cids)/20))

# Set up parallel processing: use all available cores minus 2 to avoid system overload
plan(multisession, workers = availableCores() - 2)

# Execute parallel batch queries with real-time progress tracking
with_progress({
  p <- progressor(steps = length(cid_batches))
  chem_info <- future_map_dfr(cid_batches, function(b) {
    p()
    fetch_batch(properties, b)
  })
})

# Reset to sequential processing mode
plan(sequential)

# Merge PubChem annotations with the curated compound list and standardize final output
compounds_final <- compounds_all_processed %>%
  left_join(chem_info, by = "CID") %>%
  dplyr::select(
    `Medicinal.materials(Latin.name)`,
    `Medicinal.materials(Chinese.Pinyin.name)`,
    Compound, 
    IUPACName, 
    MolecularFormula, 
    MolecularWeight, 
    CID, 
    SMILES
  ) %>%
  mutate(across(where(is.character), trimws))

save(
  compounds_final,
  file = "../data/processed/02_compounds_final.RData"
)

write.xlsx(compounds_final,"../results/02_compounds_final.xlsx")
```

## Batch Download PubChem SDF Structure Files

- Download 3D or 2D structure SDF files for all unique compounds from PubChem

```{r}
sdf_files <- "../data/raw/tcm/sdf/"

if (!dir.exists(sdf_files)) dir.create(sdf_files,showWarnings = FALSE,recursive = TRUE)

unique_cids <- unique(compounds_pubchem$CID)
results <- download_all_sdf(unique_cids, sdf_files)
```

## Reports 
```{r}
print(table(compounds_final$`Medicinal.materials(Latin.name)`))

compounds_final
```

# Acquisition of Potential targets

## Acquisition of targets from BATMAN-TCM
```{r}
rm(list = ls())

load("../data/raw/uniprot/uniprot_human_reviewed.RData") 

uniprot_human_reviewed <- uniprot_human_reviewed %>% 
  dplyr::select(Entry,Gene.Names.primary)

load("../data/processed/02_compounds_final.RData")
load("../data/raw/tcm/batman/batman_target_known.RData")
load("../data/raw/tcm/batman/batman_target_pred.RData")

# Known targets from BATMAN-TCM
compound_target_known_batman <- batman_target_known %>%
  dplyr::filter(CID %in% unique(compounds_final$CID)) %>%
  dplyr::select(CID, Uniprot.ID) %>%
  left_join(uniprot_human_reviewed, by = c("Uniprot.ID" = "Entry")) %>%
  dplyr::rename(Symbol = Gene.Names.primary) %>%
  dplyr::filter(!is.na(Symbol)) %>%
  distinct(CID, Symbol, .keep_all = TRUE) %>%
  mutate(Source = "BATMAN-Known")

# Predicted targets from BATMAN-TCM (Score ≥ 0.80,LR=48,p_adj<0.5)
compound_target_pred_batman <- batman_target_pred %>%
  dplyr::filter(CID %in% unique(compounds_final$CID)) %>%
  dplyr::select(CID, Uniprot.ID, Score) %>%
  left_join(uniprot_human_reviewed, by = c("Uniprot.ID" = "Entry")) %>%
  dplyr::rename(Symbol = Gene.Names.primary) %>%
  dplyr::filter(!is.na(Symbol), Score >= 0.80) %>%
  dplyr::select(-Score) %>%
  distinct(CID, Symbol, .keep_all = TRUE) %>%
  mutate(Source = "BATMAN-Pred")

length(unique(compound_target_pred_batman$CID))
length(unique(compound_target_pred_batman$Symbol))
length(unique(compound_target_pred_batman$Symbol)) / length(unique(compound_target_pred_batman$CID))

# Save processed BATMAN-TCM targets
save(
  compound_target_known_batman,
  compound_target_pred_batman,
  file = "../data/processed/03_batman_targets.RData"
)

```

## Acquisition of Targets from SwissTargetPrediction

```{r}
rm(list = ls())

source("utils.R")

load("../data/raw/uniprot/uniprot_human_reviewed.RData")
uniprot_human_reviewed <- uniprot_human_reviewed %>% 
  dplyr::select(Entry,Gene.Names.primary)

# Process SwissTargetPrediction results (Probability > 0.10)

compound_target_pred_swiss <- read_compound_target("../data/raw/tcm/swiss") %>%
  filter(`Probability*` > 0.10) %>%
  separate_rows(`Common name`, `Uniprot ID`, sep = " ") %>%
  mutate(
    CID = clean_cids(CID),
    Uniprot.ID = gsub("[[:space:]]", "", `Uniprot ID`)
  ) %>%
  dplyr::select(CID, Uniprot.ID) %>%
  distinct(CID, Uniprot.ID, .keep_all = TRUE) %>%
  left_join(uniprot_human_reviewed, by = c("Uniprot.ID" = "Entry")) %>%
  dplyr::rename(Symbol = Gene.Names.primary) %>%
  dplyr::filter(!is.na(Symbol)) %>%
  distinct(CID, Symbol, .keep_all = TRUE)%>%
  mutate(Source = "SwissTargetPrediction")

length(unique(compound_target_pred_swiss$Symbol))

length(unique(compound_target_pred_swiss$CID))

length(unique(compound_target_pred_swiss$Symbol)) / length(unique(compound_target_pred_swiss$CID))


# Save processed SwissTargetPrediction targets
save(
  compound_target_pred_swiss,
  file = "../data/processed/04_swiss_targets.RData"
)
```

## Acquisition of Targets from PharmMapper
```{r}
rm(list = ls())

source("utils.R")

load("../data/raw/uniprot/uniprot_human_reviewed.RData")
uniprot_human_reviewed <- uniprot_human_reviewed %>% 
  dplyr::select(Entry,Entry.Name, Gene.Names.primary)

# Process PharmMapper results (Human only, high-confidence filters) (modify,0.5)
compound_target_pred_pharm <- read_compound_target2("../data/raw/tcm/pharm",skip_row = 1) %>%
  dplyr::filter(grepl("_HUMAN", Uniplot),zscore>0, `Norm Fit`>0.5) %>%
  dplyr::select(CID, Uniplot) %>%
  distinct(CID, Uniplot, .keep_all = TRUE) %>%
  left_join(uniprot_human_reviewed, by = c("Uniplot" = "Entry.Name")) %>%
  dplyr::rename(Symbol = Gene.Names.primary) %>%
  dplyr::filter(!is.na(Symbol)) %>%
  dplyr::select(CID,Uniprot.ID = Entry,Symbol) %>%
  distinct(CID, Symbol, .keep_all = TRUE) %>%
  mutate(Source = "PharmMapper")

length(unique(compound_target_pred_pharm$Symbol))

length(unique(compound_target_pred_pharm$CID))

length(unique(compound_target_pred_pharm$Symbol)) / length(unique(compound_target_pred_pharm$CID))

# Save processed PharmMapper targets
save(
  compound_target_pred_pharm,
  file = "../data/processed/05_pharm_targets.RData"
)

```

## Acquisition of Targets from SEA
```{r}
rm(list = ls())

source("utils.R")

load("../data/raw/uniprot/uniprot_human_reviewed.RData")
uniprot_human_reviewed <- uniprot_human_reviewed %>% 
  dplyr::select(Entry, Entry.Name, Gene.Names.primary)


# Process SEA results (confidence filters)
# P < 1e-5、MaxTc >= 0.3(modify)
compound_target_pred_sea <- read.csv("../data/raw/tcm/sea/sea-results.xls") %>%
  mutate(CID = as.numeric(gsub("compound","",Query.ID)), .before = 1) %>%
  arrange(CID) %>%
  mutate(CID = as.character(CID)) %>%
  dplyr::select(-Query.ID) %>%
  dplyr::filter(
    grepl("_HUMAN", Target.ID), 
    P.Value < 1e-10, 
    Max.Tc >= 0.3
  ) %>%
  distinct(CID, Target.ID, .keep_all = TRUE) %>%
  left_join(uniprot_human_reviewed, by = c("Target.ID" = "Entry.Name")) %>%
  dplyr::rename(Symbol = Gene.Names.primary) %>%
  dplyr::filter(!is.na(Symbol)) %>%
  dplyr::select(CID,Uniprot.ID = Entry,Symbol) %>%
  distinct(CID, Symbol, .keep_all = TRUE) %>%
  mutate(Source = "SEA")

length(unique(compound_target_pred_sea$Symbol))

length(unique(compound_target_pred_sea$CID))

length(unique(compound_target_pred_sea$Symbol)) / length(unique(compound_target_pred_sea$CID))


# Save processed SEA targets
save(
  compound_target_pred_sea,
  file = "../data/processed/06_sea_targets.RData"
)

```

## Integrate All Targets and Deduplicate
```{r}
rm(list = ls())

# Load all processed target datasets
load("../data/processed/03_batman_targets.RData")
load("../data/processed/04_swiss_targets.RData")
load("../data/processed/05_pharm_targets.RData")
load("../data/processed/06_sea_targets.RData")

# Merge all targets 
compound_target <- rbind(
  compound_target_known_batman,
  compound_target_pred_batman, 
  compound_target_pred_swiss,
  compound_target_pred_pharm,
  compound_target_pred_sea
) %>%
  group_by(CID, Symbol) %>%  
  summarise(
    Source = paste(unique(na.omit(Source)), collapse = "; "),
    across(everything(), ~ dplyr::first(.)),  
    .groups = "drop"
  )

# Filter high-confidence targets (known or predicted by ≥1  databases)
compound_target_final <- compound_target %>%
  mutate(
    is_known = str_detect(Source, "Known"),
    platform_count = str_split(Source, "; ") %>%
      map_dbl(~ length(unique(.)))
  ) %>%
  filter(
    is_known | platform_count >= 1
  ) %>%
  dplyr::select(CID,Symbol,Uniprot.ID,Source)

length(unique(compound_target_final$Symbol))

length(unique(compound_target_final$CID))

length(unique(compound_target_final$Symbol)) / length(unique(compound_target_final$CID))

# Save final target list
save(
  compound_target_final,
  file = "../data/processed/07_compound_target_final.RData"
)

write.xlsx(compound_target_final, "../results/03_compound_target_final.xlsx")
```

## Target Statistics Across Prediction Platforms.
```{r}
# Generate summary table for manuscript
summary_table <- tibble(
  Database = c(
    "BATMAN-TCM (Known)",
    "BATMAN-TCM (Predicted)",
    "SwissTargetPrediction",
    "PharmMapper",
    "SEA",
    "Final Integrated Results"
  ),
  
  # Number of unique compounds with targets (CID)
  `Number of compounds with targets` = c(
    n_distinct(compound_target_known_batman$CID),
    n_distinct(compound_target_pred_batman$CID),
    n_distinct(compound_target_pred_swiss$CID),
    n_distinct(compound_target_pred_pharm$CID),
    n_distinct(compound_target_pred_sea$CID),
    n_distinct(compound_target_final$CID)
  ),
  
  # Number of unique targets based on UniProt ID
  `Unique Targets (UniProt ID)` = c(
    n_distinct(compound_target_known_batman$Uniprot.ID),
    n_distinct(compound_target_pred_batman$Uniprot.ID),
    n_distinct(compound_target_pred_swiss$Uniprot.ID),
    n_distinct(compound_target_pred_pharm$Uniprot.ID),
    n_distinct(compound_target_pred_sea$Uniprot.ID),
    n_distinct(compound_target_final$Uniprot.ID)
  )
)

# Save statistics

save(summary_table, file =  "../data/processed/08_target_statistics.RData")
```

## herb-compound-target
```{r}
rm(list = ls())

load("../data/processed/02_compounds_final.RData")
load("../data/processed/07_compound_target_final.RData")

herb_compound_target <- compound_target_final %>%
  left_join(
    compounds_final,
    by = "CID",
    relationship = "many-to-many"
  ) %>%
  dplyr::select(
    `Medicinal.materials(Latin.name)`, 
    `Medicinal.materials(Chinese.Pinyin.name)`, 
    Compound, 
    CID, 
    MolecularFormula, 
    MolecularWeight, 
    SMILES, 
    IUPACName,
    Symbol, 
    Uniprot.ID
  ) %>%
  distinct(`Medicinal.materials(Latin.name)`, CID, Symbol, .keep_all = TRUE)

length(unique(herb_compound_target$`Medicinal.materials(Latin.name)`))
length(unique(herb_compound_target$CID))
length(unique(herb_compound_target$Symbol))

save(herb_compound_target, file = "../data/processed/09_herb_compound_target.RData")
```
