---
title: "原始数据预处理"

execute:
  eval: false
  echo: true
  warning: false
  message: false
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = TRUE,
  collapse = TRUE,
  comment = "#>",
  cache = FALSE,
  class.output="scroll-100",
  cache.path = "cached/",
  fig.width=8, 
  fig.height=6, 
  dpi=300)

library(tidyverse)
library(openxlsx)
library(data.table)
library(rmdwc)
# remotes::install_github("sigbertklinke/rmdwc", dependencies = TRUE)

library(PubChemR)
library(future)     
library(furrr)      
library(progressr) 
library(httr) 
library(progress)
library(jsonlite)

# https://www.uniprot.org/id-mapping

# library(biomaRt)
# library(clusterProfiler)
# library(org.Hs.eg.db)

# library(rentrez)
# install.packages("rentrez")

# library(mygene)
# BiocManager::install("mygene")

library(disgenet2r)
api_key <- "b36b7049-f17d-4444-af51-e23b2262b355"
# api_key <- "enter your API key here"
Sys.setenv(DISGENET_API_KEY= api_key)
# https://gitlab.com/medbio/disgenet2r
```

```{css css_options, echo=FALSE}
pre {
  max-height: 300px;      
  overflow-y: auto;       
}
```

```{css css_scroll, echo=FALSE}
.scroll-100 {
  max-height: 100px;     
  overflow-y: auto;       
  background-color: inherit; 
}
```


---

# 活性成分

- 原始数据：根据文献[@zhang2024qfd; @tan2012qfda]获得了 AR 和 ALRP 的入血活性成分信息（../data/qfd_ingredients/raw.xlsx）

- 输出数据：(supplementary/Table S1.xlsx)

## 从PubChem获取分子信息
```{r}
rm(list = ls())

# 读取原始成分数据
ingredients_raw <- read.xlsx("../data/qfd_ingredients/raw.xlsx") |>
  mutate(PubChem.CID = as.character(PubChem.CID)) |>
  distinct(Abbreviation, PubChem.CID, .keep_all = TRUE) |>
  arrange(Abbreviation)

properties <- c("IUPACName", "MolecularFormula", "MolecularWeight", "SMILES","InChIKey")
unique_cids <- unique(ingredients_raw$PubChem.CID)

chem_info <- get_properties(
  properties = properties,
  identifier = unique_cids,
  namespace = "cid" 
) |> 
  retrieve(., .combine.all = TRUE)

ingredients <- ingredients_raw |>
  mutate(PubChem.CID = as.numeric(PubChem.CID)) |>
  left_join(chem_info, by = c("PubChem.CID"="CID")) |>
  mutate(across(where(is.character), trimws)) |>
  mutate(MolecularWeight = as.numeric(MolecularWeight)) |>
  arrange(Abbreviation, MolecularWeight)

save(ingredients,file = "../data/qfd_ingredients/ingredients.RData")
```

## 保存分子信息到Table S1
```{r}
rm(list = ls())

load("../data/qfd_ingredients/ingredients.RData")

write.xlsx(list(
  `Table S1` = ingredients |>
    dplyr::rename(`Medicinal Materials` = Medicinal.Materials,
                `Molecular Formula` = MolecularFormula,
                `PubChem CID` = PubChem.CID,
                `TMMC2 ID` = TMMC2.ID,
                `HERB2 ID` = HERB2.ID,
                `TCMSP ID` = TCMSP.ID,
                `Molecular Weight` = MolecularWeight,
                `IUPAC Name` = IUPACName
                ) |>
    dplyr::select(`Medicinal Materials`, Abbreviation, Ingredient,`PubChem CID`, InChIKey,`Molecular Formula`,`Molecular Weight`, SMILES,ConnectivitySMILES)), 
  file = "supplementary/Table S1.xlsx")
```

## 下载SDF文件(../data/qfd_sdfs)
```{r}
rm(list = ls())

load("../data/qfd_ingredients/ingredients.RData")

unique_cids <- unique(ingredients$PubChem.CID)
sdf_files <- "../data/qfd_sdfs"

source("utils.R")
results <- download_all_sdf(unique_cids, sdf_files)

# output_dir <- "../data/qfd_sdfs2"    
# 
# if (!dir.exists(output_dir)) {
#   dir.create(output_dir, recursive = TRUE)
# }
# 
# plan(multisession, workers = 4)
# future_walk(.x = unique_cids, .f = ~ {
#   get_sdf(
#     identifier = .x,
#     namespace = "cid",
#     path = output_dir,
#     file_name = as.character(.x),
#     options = list(record_type = "3d")
#   )
#   Sys.sleep(0.3)
# }, .progress = TRUE)  
# plan(sequential)
# 
# list.files(output_dir, pattern = "\\.sdf$")
```

# 成分靶点

- 根据成分信息(名称, 分子式, CID, SMILES)从`PubChem BioAssay`, `STITCH v5.0`,`TCMSP v2.3`, `HERB2.0`，  `SwissTargetPrediction`, `Super-PRED`, `PharmMapper`,  `SEA`等8个靶点数据库或预测平台获得了成分的潜在靶点(supplementary/Table S2.xlsx)

- 使用`UniProt`进行标准化

## TMMC 2.0
```{r}
rm(list = ls())

load("../data/qfd_targets/tmmc2/tmmc.RData")
load("../data/qfd_ingredients/ingredients.RData")

targets.tmmc <- tmmc_compound_protein |>
  dplyr::filter(ID %in% ingredients$TMMC2.ID) |>
  arrange(ID)

save(targets.tmmc, file = "../data/qfd_targets/tmmc2/targets.tmmc.RData")
```

## PubChem BioAssay

### gene.symbol
```{r}
rm(list = ls())

# source("utils.R")

load("../data/qfd_ingredients/ingredients.RData")

# load("../data/qfd_targets/tmmc2/targets.tmmc.RData")
# targets.pubchem <- targets.tmmc |>
#   dplyr::filter(SOURCE == "PubChem") |>
#   dplyr::distinct(ID,PROTEINID, .keep_all = TRUE) |>
#   left_join(ingredients, by =  c("ID" = "TMMC2.ID"), relationship = "many-to-many") |>
#   mutate(Class = "validated") |>
#   dplyr::select(No.,Abbreviation,Ingredient,Type,PubChem.CID, PROTEINID, Source = SOURCE, Score = SCORE, Class) |>
#   distinct(Abbreviation,PubChem.CID,PROTEINID, Source, Class, .keep_all = TRUE) |>
#   dplyr::filter(if_all(everything(), ~ !is.na(.x))) |>
#   arrange(No., PROTEINID)

source("pubchemBioassay.R")

cids <- as.character(ingredients$PubChem.CID)

# Chemical-Target Interactions

targets_cti <- get_cid_targets(cids) 

save(targets_cti, file = "../data/qfd_targets/pubchem/targets_cti.RData")

targets.pubchem.interaction <- targets_cti |>
  dplyr::filter(taxid == "9606") |>
  dplyr::select(CID = cid, Symbol = genename) |>
  distinct(CID,Symbol) |>
  mutate(Source = "PubChem.Interaction", Class="validated")

save(targets.pubchem.interaction, file = "../data/qfd_targets/pubchem/targets.pubchem.interaction.RData")
  

# Bioassay
# active_aids <- get_cid_aid_df(cids)

# unique_active_aids <- unique(active_aids$AID)

bioassay <- get_assays_by_cids(cids)

save(bioassay, file = "../data/qfd_targets/pubchem/bioassay.RData")

bioassay_active_targets <- bioassay |>
  dplyr::filter(`Activity Outcome` == "Active") |>
  dplyr::filter(`Target GeneID` !="" | `Target Accession` != "") 

save(bioassay_active_targets, file = "../data/qfd_targets/pubchem/bioassay_active_targets.RData")

gene.symbol.by.geneid <- get_gene_symbol_by_geneids(bioassay_active_targets$`Target GeneID`) |>
  dplyr::filter((!is.na(Symbol)) & TaxonomyID == "9606")

save(gene.symbol.by.geneid, file = "../data/qfd_targets/pubchem/gene.symbol.by.geneid.RData")

accession_df <- bioassay_active_targets |>
  dplyr::filter(`Target Accession`!="" & `Target GeneID` =="") |>
  dplyr::select(Accession = `Target Accession`) |>
  distinct(Accession)

gene.symbol.by.accession <- get_gene_symbol_by_accessions(accession_df$Accession) |>
  dplyr::filter((!is.na(Symbol)) & TaxonomyID == "9606")

save(gene.symbol.by.accession, file = "../data/qfd_targets/pubchem/gene.symbol.by.accession.RData")

gene.pubchem.bioassay.geneid <- bioassay_active_targets |>
  left_join(gene.symbol.by.geneid, 
            by = c("Target GeneID" = "GeneID"),
            relationship = "many-to-many") |>
  dplyr::filter(!is.na(Symbol))

gene.pubchem.bioassay.accession <- bioassay_active_targets |>
  left_join(gene.symbol.by.accession, 
            by = c("Target Accession" = "Accession"),
            relationship = "many-to-many") |>
  dplyr::filter(!is.na(Symbol))

gene.pubchem.bioassay <- rbind(
  gene.pubchem.bioassay.geneid,
  gene.pubchem.bioassay.accession
) |>
  distinct(CID, Symbol) |>
  mutate(Source = "PubChem.Interaction", Class="validated") 

save(gene.pubchem.bioassay.geneid,
     gene.pubchem.bioassay.accession,
     gene.pubchem.bioassay,
     file = "../data/qfd_targets/pubchem/gene.pubchem.bioassay.accession.RData")
  

gene.pubchem <- rbind(
  gene.pubchem.interaction,
  gene.pubchem.bioassay
) |>
  distinct(CID,Symbol) |>
  mutate(Source = "PubChem", Class="validated") 


save(gene.pubchem, file = "../data/qfd_targets/pubchem/gene.pubchem.RData")
```

### targets.pubchem
```{r}
rm(list = ls())
load("../data/qfd_targets/pubchem/gene.pubchem.RData")
load("../data/qfd_ingredients/ingredients.RData")

targets.pubchem <- ingredients |>
  mutate(PubChem.CID = as.character(PubChem.CID)) |>
  left_join(gene.pubchem, by = c("PubChem.CID"="CID"), relationship = "many-to-many") |>
  dplyr::filter(!is.na(Symbol)) |>
  dplyr::select(No.,Abbreviation,Ingredient,Type,PubChem.CID,Symbol, Source, Class) |>
  distinct(Abbreviation,PubChem.CID,Symbol, Source, Class, .keep_all = TRUE) |>
  dplyr::filter(if_all(everything(), ~ !is.na(.x))) |>
  arrange(No., PubChem.CID)

save(targets.pubchem, file = "../data/qfd_targets/pubchem/targets.pubchem.RData")
``` 

## STITCH v5.0 (score>=0.4)

```{r}
rm(list = ls())

source("utils.R")

load("../data/qfd_ingredients/ingredients.RData")

stitch <- read_compound_target("../data/qfd_targets/stitch5/") |>
  dplyr::filter(node1_string_internal_id > 0 & node2_string_internal_id <0 ) |>
  arrange(node2) |>
  mutate(
    Source = "STITCH V5.0",
    Class = "predicted"
  )

targets.stitch <- ingredients  |>
  mutate(PubChem.CID = as.character(PubChem.CID)) |>
  left_join(stitch, by = c("PubChem.CID"="CID"), relationship = "many-to-many") |>
  dplyr::filter(!is.na(`#node1`)) |>
  dplyr::select(No.,Abbreviation,Ingredient,Type,PubChem.CID,Symbol = `#node1`, Source, Score = combined_score, Class) |>
  distinct(Abbreviation,PubChem.CID,Symbol, Source, Class, .keep_all = TRUE) |>
  dplyr::filter(if_all(everything(), ~ !is.na(.x))) |>
  arrange(No., PubChem.CID)

save(targets.stitch, file = "../data/qfd_targets/stitch5/targets.stitch.RData")

# 
# rm(list = ls())
# 
# load("../data/qfd_targets/tmmc2/targets.tmmc.RData")
# load("../data/qfd_ingredients/ingredients.RData")
# 
# targets.stitch <- targets.tmmc |>
#   dplyr::filter(SOURCE == "STITCH") |>
#   dplyr::distinct(ID,PREFERRED_NAME, .keep_all = TRUE) |>
#   left_join(ingredients, by =  c("ID" = "TMMC2.ID"), relationship = "many-to-many") |>
#   mutate(Class = "predicted") |>
#   dplyr::select(No.,Abbreviation,Ingredient,Type,PubChem.CID, PREFERRED_NAME, Source = SOURCE, Score = SCORE, Class) |>
#   dplyr::filter(Score >=0.4) |>
#   distinct(Abbreviation,PubChem.CID,PREFERRED_NAME, Source, Class, .keep_all = TRUE) |>
#   dplyr::filter(if_all(everything(), ~ !is.na(.x))) |>
#   arrange(No., Ingredient) 
```

## TCMSP v2.3

```{r}
rm(list = ls())

load("../data/qfd_ingredients/ingredients.RData")

tcmsp <- read.xlsx("../data/qfd_targets/tcmsp/tcmsp_qfd_targets.xlsx") |>
  dplyr::mutate(
    Source = "TCMSP",
    Class = case_when(
    validated !="validated" ~ "predicted",
    TRUE ~ "validated"),
    Score.SVM = SVM_score
  ) |>
  dplyr::select(MOL_ID,Entry,Source, Score.SVM, Class) 

targets.tcmsp <- ingredients  |>
  separate_rows(TCMSP.ID, sep = ";") |>
  mutate(TCMSP.ID = trimws(TCMSP.ID)) |>
  left_join(tcmsp, by = c("TCMSP.ID" = "MOL_ID"), relationship = "many-to-many") |>
  dplyr::select(No.,Abbreviation,Ingredient,Type,PubChem.CID,Entry, Source, Score.SVM, Class) |>
  distinct(Abbreviation,PubChem.CID,Entry, Source, Class, .keep_all = TRUE) |>
  dplyr::filter(if_all(everything(), ~ !is.na(.x))) |>
  arrange(No., Entry)

save(targets.tcmsp, file = "../data/qfd_targets/tcmsp/targets.tcmsp.RData")

```

## HERB 2.0
```{r}
rm(list = ls())

load("../data/qfd_ingredients/ingredients.RData")

herb2_db_all <- read.xlsx("../data/qfd_targets/herb2/herb2_db_all.xlsx") |>
  mutate(Source = "HERB2", Mining = "database", Class = "validated") |>
  dplyr::select(PubChem.CID, Symbol = Gene.symbol, Source,Mining, Class)

herb2_ref_all <- read.xlsx("../data/qfd_targets/herb2/herb2_ref_all.xlsx") |>
  mutate(Source = "HERB2", Mining = "Reference", Class = "validated") |>
  dplyr::select(PubChem.CID, Symbol = Gene.symbol, Source,Mining, Class)


herb2 <- rbind(herb2_db_all, herb2_ref_all) |>
  distinct(PubChem.CID,Symbol, Mining, Class, .keep_all = TRUE)

targets.herb2 <- ingredients |>
  mutate(PubChem.CID = as.character(PubChem.CID)) |>
  left_join(herb2, by = "PubChem.CID", relationship = "many-to-many") |>
  dplyr::select(No.,Abbreviation,Ingredient,Type,PubChem.CID,Symbol, Source, Mining, Class) |>
  distinct(Abbreviation,PubChem.CID,Symbol, Source, Class, .keep_all = TRUE) |>
  dplyr::filter(if_all(everything(), ~ !is.na(.x))) |>
  arrange(No., Symbol)

save(targets.herb2, file = "../data/qfd_targets/herb2/targets.herb2.RData")
```

## SwissTargetPrediction

```{r}
rm(list = ls())

source("utils.R")

load("../data/qfd_ingredients/ingredients.RData")

swiss <- read_compound_target("../data/qfd_targets/swiss/pred") |>
  dplyr::rename(Probability = `Probability*`, 
                Common.name = `Common name`,
                Uniprot.ID = `Uniprot ID`) |>
  separate_rows(Uniprot.ID, sep = "&") |>
  separate_rows(Uniprot.ID, sep = " ") |>
  dplyr::select(CID,Uniprot.ID, Probability) |>
  distinct(CID, Uniprot.ID, .keep_all = TRUE) |>
  dplyr::filter(!is.na(Uniprot.ID)) |>
  distinct(CID, Uniprot.ID, .keep_all = TRUE) |>
  mutate(
    CID = clean_cids(CID),
    Source = "SwissTargetPrediction",
    Class = "predicted"
    ) |>
  dplyr::select(CID, Uniprot.ID,Source,Probability,Class)

targets.swiss <- ingredients |>
  mutate(PubChem.CID = as.character(PubChem.CID)) |>
  left_join(swiss, by = c("PubChem.CID" = "CID"), relationship = "many-to-many") |>
  dplyr::select(No.,Abbreviation,Ingredient,Type,PubChem.CID,Uniprot.ID, Source, Probability, Class) |>
  distinct(Abbreviation,PubChem.CID, Uniprot.ID, Source, Class, .keep_all = TRUE) |>
  dplyr::filter(if_all(everything(), ~ !is.na(.x))) |>
  arrange(No., Uniprot.ID)

# Save processed SwissTargetPrediction targets
save(targets.swiss, file = "../data/qfd_targets/swiss/targets.swiss.RData")

```

## Super-PRED
```{r}
rm(list = ls())

source("utils.R")

load("../data/qfd_ingredients/ingredients.RData")

# known targets
super_known <- read_compound_target("../data/qfd_targets/super/known") |>
  dplyr::select(CID, Uniprot.ID = `UniProt ID`) |>
  dplyr::filter(!is.na(Uniprot.ID)) |>
  distinct(CID, Uniprot.ID, .keep_all = TRUE) |>
  mutate(Source = "SuperPred", Probability = "100%", `Model accuracy` = "100%", Class = "validated") |>
  dplyr::select(CID, Uniprot.ID, Source, Probability, `Model accuracy`, Class)

# Predicted targets (Prob >= 60, Model Accuracy >= 85) (modify,70/90)
super_pred <- read_compound_target("../data/qfd_targets/super/pred") |>
  dplyr::select(CID, Uniprot.ID = `UniProt ID`, Probability, `Model accuracy`) |>
  dplyr::filter(!is.na(Uniprot.ID)) |>
  distinct(CID, Uniprot.ID, .keep_all = TRUE)|>
  mutate(Source = "SuperPred", Class = "predicted") |>
  dplyr::select(CID, Uniprot.ID, Source, Probability, `Model accuracy`, Class)

super <- rbind(super_known, super_pred)

targets.super <- ingredients |>
  mutate(PubChem.CID = as.character(PubChem.CID)) |>
  left_join(super, by = c("PubChem.CID" = "CID"), relationship = "many-to-many") |>
  dplyr::select(No.,Abbreviation,Ingredient,Type,PubChem.CID,Uniprot.ID, Source, Probability, `Model accuracy`, Class) |>
  distinct(Abbreviation,PubChem.CID, Uniprot.ID, Source, Class, .keep_all = TRUE) |>
  dplyr::filter(if_all(everything(), ~ !is.na(.x))) |>
  arrange(No., Uniprot.ID)

# Save processed Super-Pred targets
save(targets.super, file = "../data/qfd_targets/super/targets.super.RData")

```

## PharmMapper

```{r}
rm(list = ls())

source("utils.R")

load("../data/qfd_ingredients/ingredients.RData")

# Process PharmMapper results (Human only, high-confidence filters) (modify,0.5)
pharm <- read_compound_target("../data/qfd_targets/pharm/pred",skip_row = 1) |>
  dplyr::filter(grepl("_HUMAN", Uniplot)) |>
  dplyr::select(CID, Uniplot,`Norm Fit`) |>
  distinct(CID, Uniplot, .keep_all = TRUE) |>
  dplyr::filter(!is.na(Uniplot)) |>
  mutate(Source = "PharmMapper",Class = "predicted") |>
  dplyr::select(CID,Uniplot, Source,`Norm Fit`,Class) |>
  distinct(CID, Uniplot, .keep_all = TRUE) 

targets.pharm <- ingredients |>
  mutate(PubChem.CID = as.character(PubChem.CID)) |>
  left_join(pharm, by = c("PubChem.CID" = "CID"), relationship = "many-to-many") |>
  dplyr::select(No.,Abbreviation,Ingredient,Type,PubChem.CID,Uniplot, Source,`Norm Fit`, Class) |>
  distinct(Abbreviation,PubChem.CID, Uniplot, Source, Class, .keep_all = TRUE) |>
  dplyr::filter(if_all(everything(), ~ !is.na(.x))) |>
  arrange(No., Uniplot)

# Save processed PharmMapper targets
save(targets.pharm, file = "../data/qfd_targets/pharm/targets.pharm.RData")
```

## SEA

```{r}
rm(list = ls())

source("utils.R")

load("../data/qfd_ingredients/ingredients.RData")

# Process SEA results (confidence filters)
sea <- read.csv("../data/qfd_targets/sea/sea-results.xls") |>
  mutate(CID = as.numeric(gsub("C","",Query.ID)), .before = 1) |>
  arrange(CID) |>
  mutate(CID = as.character(CID)) |>
  dplyr::select(-Query.ID) |>
  dplyr::filter(grepl("_HUMAN", Target.ID)) |>
  distinct(CID, Target.ID, .keep_all = TRUE) |>
  dplyr::filter(!is.na(Target.ID)) |>
  mutate(Source = "SEA", Class = "predicted") |>
  dplyr::select(CID, Target.ID, Source, P.Value, Max.Tc, Class) |>
  distinct(CID, Target.ID, .keep_all = TRUE) 

targets.sea <- ingredients |>
  mutate(PubChem.CID = as.character(PubChem.CID)) |>
  left_join(sea, by = c("PubChem.CID" = "CID"), relationship = "many-to-many") |>
  dplyr::select(No.,Abbreviation,Ingredient,Type,PubChem.CID,Target.ID, Source, P.Value, Max.Tc,Class) |>
  distinct(Abbreviation,PubChem.CID, Target.ID, Source, Class, .keep_all = TRUE) |>
  dplyr::filter(if_all(everything(), ~ !is.na(.x))) |>
  arrange(No., Target.ID)

# Save processed SEA targets
save(targets.sea, file = "../data/qfd_targets/sea/targets.sea.RData")
```


## 筛选潜在成分靶点

```{r}
rm(list = ls())

# pubchem: all
load("../data/qfd_targets/pubchem/targets.pubchem.RData")
pubchem_filter <- targets.pubchem

save(pubchem_filter, file = "../data/qfd_targets/pubchem/pubchem_filter.RData")
write.xlsx(
  pubchem_filter |> 
    distinct(Symbol),
  file = "../data/qfd_targets/pubchem/pubchem_filter.xlsx"
)


# stitch: Score > 0.7
load("../data/qfd_targets/stitch5/targets.stitch.RData")
stitch_filter <- targets.stitch |>
  dplyr::filter(Score>=0.7) |>
  dplyr::select(-Score)

save(stitch_filter, file = "../data/qfd_targets/stitch5/stitch_filter.RData")
write.xlsx(
  stitch_filter |> 
    distinct(Symbol),
  file = "../data/qfd_targets/stitch5/stitch_filter.xlsx"
)


# TCMSP: all
load("../data/qfd_targets/tcmsp/targets.tcmsp.RData")
tcmsp_filter <- targets.tcmsp |> dplyr::select(-Score.SVM)

save(tcmsp_filter, file = "../data/qfd_targets/tcmsp/tcmsp_filter.RData")
write.xlsx(
  tcmsp_filter |> 
    distinct(Entry),
  file = "../data/qfd_targets/tcmsp/tcmsp_filter.xlsx"
)


# HERB2: all
load("../data/qfd_targets/herb2/targets.herb2.RData")
herb2_filter <- targets.herb2 |> dplyr::select(-Mining)

save(herb2_filter, file = "../data/qfd_targets/herb2/herb2_filter.RData")
write.xlsx(
  herb2_filter |> 
    distinct(Symbol),
  file = "../data/qfd_targets/herb2/herb2_filter.xlsx"
)

# Swiss: Probability>=0.1
load("../data/qfd_targets/swiss/targets.swiss.RData")
swiss_filter <- targets.swiss |>
  dplyr::filter(Probability>=0.1) |>
  dplyr::select(-Probability)

save(swiss_filter, file = "../data/qfd_targets/swiss/swiss_filter.RData")

write.xlsx(
  swiss_filter |> 
    distinct(Uniprot.ID),
  file = "../data/qfd_targets/swiss/swiss_filter.xlsx"
)


# Super-PRED: Probability>="80%" (+ `Model accuracy` >= "95%")
load("../data/qfd_targets/super/targets.super.RData")
super_filter <- targets.super |>
  mutate(Prob_num = as.numeric(str_remove(Probability,"%"))) |>
  dplyr::filter((Prob_num>=80 ) | Class == "validated") |>
  dplyr::select(-Probability,-`Model accuracy`, -Prob_num)

save(super_filter, file = "../data/qfd_targets/super/super_filter.RData")
write.xlsx(
  super_filter |> 
    distinct(Uniprot.ID),
  file = "../data/qfd_targets/super/super_filter.xlsx"
)


# PharmMapper: `Norm Fit` >= 0.70
load("../data/qfd_targets/pharm/targets.pharm.RData")
pharm_filter <- targets.pharm |> 
  dplyr::filter(`Norm Fit` >= 0.7) |>
  dplyr::select(-`Norm Fit`)

save(pharm_filter, file = "../data/qfd_targets/pharm/pharm_filter.RData")
write.xlsx(
  pharm_filter |> 
    distinct(Uniplot),
  file = "../data/qfd_targets/pharm/pharm_filter.xlsx"
)

# SEA: P.Value < 0.01、Max.Tc >= 0.3
load("../data/qfd_targets/sea/targets.sea.RData")
sea_filter <- targets.sea |> 
  dplyr::filter(P.Value < 0.01 & Max.Tc >= 0.3) |>
  dplyr::select(-P.Value, -Max.Tc)

save(sea_filter, file = "../data/qfd_targets/sea/sea_filter.RData")

write.xlsx(
  sea_filter |> 
    distinct(Target.ID),
  file = "../data/qfd_targets/sea/sea_filter.xlsx"
)

```


## Uniprot ID Mapping

<!-- - pubchem_mapping.tsv.gz -->
<!-- - stitch_mapping.tsv.gz -->
<!-- - tcmsp_mapping.tsv.gz -->
<!-- - herb2_mapping.tsv.gz -->
- swiss_mapping.tsv.gz
- super_mapping.tsv.gz
- pharm_mapping.tsv.gz
- sea_mapping.tsv.gz


```{r}
rm(list = ls())

# pubchem
load("../data/qfd_targets/pubchem/pubchem_filter.RData")
pubchem_mapping <- pubchem_filter |>
  dplyr::select(
    Abbreviation, 
    Ingredient, 
    PubChem.CID, 
    Symbol, 
    Source,
    Class
  )

save(pubchem_mapping, file = "../data/qfd_targets/pubchem/pubchem_mapping.RData")

# stitch
load("../data/qfd_targets/stitch5/stitch_filter.RData")
stitch_mapping <- stitch_filter |>
  dplyr::select(
    Abbreviation, 
    Ingredient, 
    PubChem.CID, 
    Symbol, 
    Source,
    Class
  )

save(stitch_mapping, file = "../data/qfd_targets/stitch5/stitch_mapping.RData")


# tcmsp
load("../data/qfd_targets/tcmsp/tcmsp_filter.RData")
tcmsp_mapping <-
  fread("../data/qfd_targets/tcmsp/tcmsp_mapping.tsv.gz") |>
  dplyr::rename(Symbol = To) |>
  right_join(tcmsp_filter, by = c("From" = "Entry"),
            relationship = "many-to-many") |>
  dplyr::filter(!is.na(Symbol)) |>
  distinct(Abbreviation,PubChem.CID,Symbol,.keep_all = TRUE) |>
  dplyr::select(
    Abbreviation, 
    Ingredient, 
    PubChem.CID, 
    Symbol, 
    Source,
    Class
  )

save(tcmsp_mapping, file = "../data/qfd_targets/tcmsp/tcmsp_mapping.RData")

# herb
load("../data/qfd_targets/herb2/herb2_filter.RData")
herb2_mapping <- herb2_filter |>
  dplyr::select(
    Abbreviation, 
    Ingredient, 
    PubChem.CID, 
    Symbol, 
    Source,
    Class
  ) 
save(herb2_mapping, file = "../data/qfd_targets/herb2/herb2_mapping.RData")

# swiss
load("../data/qfd_targets/swiss/swiss_filter.RData")
swiss_mapping <- fread("../data/qfd_targets/swiss/swiss_mapping.tsv.gz") |>
  dplyr::rename(Symbol = To) |>
  distinct(From, Symbol) |>
  right_join(swiss_filter, by = c("From" = "Uniprot.ID"),
            relationship = "many-to-many") |>
  dplyr::filter(!is.na(Symbol)) |>
  distinct(Abbreviation,PubChem.CID,Symbol,.keep_all = TRUE) |>
  dplyr::select(
    Abbreviation, 
    Ingredient, 
    PubChem.CID, 
    Symbol, 
    Source,
    Class
  ) 
save(swiss_mapping, file = "../data/qfd_targets/swiss/swiss_mapping.RData")

# super
load("../data/qfd_targets/super/super_filter.RData")
super_mapping <- fread("../data/qfd_targets/super/super_mapping.tsv.gz") |>
  dplyr::rename(Symbol = To) |>
  distinct(From, Symbol) |>
  separate_rows(Symbol, sep = ";") |>
  right_join(super_filter, by = c("From" = "Uniprot.ID"),
            relationship = "many-to-many") |>
  dplyr::filter(!is.na(Symbol)) |>
  distinct(Abbreviation,PubChem.CID,Symbol,.keep_all = TRUE) |>
  dplyr::select(
    Abbreviation, 
    Ingredient, 
    PubChem.CID, 
    Symbol, 
    Source,
    Class
  ) 
save(super_mapping, file = "../data/qfd_targets/super/super_mapping.RData")

# pharm
load("../data/qfd_targets/pharm/pharm_filter.RData")
pharm_mapping <- fread("../data/qfd_targets/pharm/pharm_mapping.tsv.gz") |>
  dplyr::rename(Symbol = To) |>
  distinct(From, Symbol) |>
  separate_rows(Symbol, sep = ";") |>
  left_join(pharm_filter, by = c("From" = "Uniplot"),
            relationship = "many-to-many") |>
  dplyr::filter(!is.na(Symbol)) |>
  distinct(Abbreviation,PubChem.CID,Symbol,.keep_all = TRUE) |>
  dplyr::select(
    Abbreviation, 
    Ingredient, 
    PubChem.CID, 
    Symbol, 
    Source,
    Class
  ) 
save(pharm_mapping, file = "../data/qfd_targets/pharm/pharm_mapping.RData")

# sea
load("../data/qfd_targets/sea/sea_filter.RData")
sea_mapping <- fread("../data/qfd_targets/sea/sea_mapping.tsv.gz") |>
  dplyr::rename(Symbol = To) |>
  distinct(From, Symbol) |>
  separate_rows(Symbol, sep = ";") |>
  right_join(sea_filter, by = c("From"= "Target.ID" ),
            relationship = "many-to-many") |>
  dplyr::filter(!is.na(Symbol)) |>
  distinct(Abbreviation,PubChem.CID,Symbol,.keep_all = TRUE) |>
  dplyr::select(
    Abbreviation, 
    Ingredient, 
    PubChem.CID, 
    Symbol, 
    Source,
    Class
  ) 
save(sea_mapping, file = "../data/qfd_targets/sea/sea_mapping.RData")

save(
  pubchem_mapping,
  stitch_mapping,
  tcmsp_mapping,
  herb2_mapping,
  swiss_mapping,
  super_mapping,
  pharm_mapping,
  sea_mapping,
  file = "../data/qfd_targets/ingredients.mapping.RData"
)
```


## 合并潜在成分靶点
```{r}
rm(list = ls())
load("../data/qfd_targets/ingredients.mapping.RData")

ingredients.targets <- rbind(
  as.data.frame(pubchem_mapping),
  as.data.frame(stitch_mapping),
  as.data.frame(tcmsp_mapping),
  as.data.frame(herb2_mapping),
  as.data.frame(swiss_mapping),
  as.data.frame(super_mapping),
  as.data.frame(pharm_mapping),
  as.data.frame(sea_mapping)
) |> 
  group_by(Abbreviation, Ingredient, PubChem.CID, Symbol) |>
  summarise(
    Source.Sum = paste(Source, collapse = ";"),  
    Class.Sum = paste(Class, collapse = ";"), 
    Count = n_distinct(Source),      
    .groups = "drop"
  ) |>
  mutate(Class = case_when(
    str_detect(Class.Sum, "validated") ~ "validated",
    TRUE ~ "predicted"
  )) |>
  mutate(Symbol = trimws(Symbol))

ingredients.targets.filter <-  ingredients.targets |>
  # dplyr::filter(Class == "validated" | (Class == "predicted" & Count>=2))
  dplyr::filter(
    (Abbreviation == "ALRP") |  # 附子全部保留，不做任何过滤
    (Abbreviation == "AR" & (Class == "validated" | (Class == "predicted" & Count >= 2)))
  )

ingredients.targets.alrp <- ingredients.targets.filter |>
  dplyr::filter(Abbreviation == "ALRP")
cat("\n附子潜在靶点数：", length(unique(ingredients.targets.alrp$Symbol)))

ingredients.targets.alrp <- ingredients.targets.filter |>
  dplyr::filter(Abbreviation == "AR")
cat("\n黄芪潜在靶点数：", length(unique(ingredients.targets.alrp$Symbol)))

cat("\n潜在靶点总数：", length(unique(ingredients.targets.filter$Symbol)))


save(ingredients.targets,ingredients.targets.filter, file =  "../data/qfd_targets/ingredients_targets.RData")

write.xlsx(ingredients.targets.filter, file = "supplementary/Table S2.xlsx")
```

# 疾病靶点

"chronic heart failure", "heart failure"

## DisGeNet v25.4 (2025-12-15更新/2025-12-20检索)
```{r}
rm(list = ls())

res <- disease2gene(disease = "UMLS_C0018801", 
                    database = "CURATED",
                    score = c(0,1) 
)

hf <- res@qresult

res <- disease2gene(disease = "UMLS_C0018802",
                    database = "CURATED",
                    score = c(0,1) 
)

chf <- res@qresult

load("../data/uniprot/uniprot_human_reviewed.RData")
uniprot_human_reviewed <- uniprot_human_reviewed |>
  mutate(Gene.Names.primary = trimws(Gene.Names.primary),
         Gene.Names = trimws(Gene.Names)) |>
  dplyr::filter(Gene.Names.primary != "") |>
  dplyr::select(Gene.Names.primary,Gene.Names) |>
  separate_rows(Gene.Names, sep = "\\s+") |>
  dplyr::filter(Gene.Names !="") |>
  mutate(Gene.Names.primary = trimws(Gene.Names.primary),
         Gene.Names = trimws(Gene.Names))

disgenet <- rbind(hf,chf) |>
  filter(geneNcbiType == "protein-coding") |>
  group_by(gene_symbol) |>
  summarise(
    max_normalized_score = max(normalized_score, na.rm = TRUE),
    Count = n_distinct(disease_name),
    .groups = "drop"
  ) |>
  left_join(uniprot_human_reviewed, 
            by = c("gene_symbol" = "Gene.Names"),
            relationship = "many-to-many") |>
  dplyr::filter(!is.na(Gene.Names.primary)) |>
  distinct(Gene.Names.primary, .keep_all = TRUE) |>
  dplyr::rename(Symbol = Gene.Names.primary) |>
  mutate(Source = "DisGeNet V25.4") |>
  dplyr::select(Symbol,Source,max_normalized_score,Count)

nrow(disgenet)

save(hf,chf, file = "../data/chf_targets/raw.RData")

save(disgenet,file = "../data/chf_targets/disgenet.RData")

write.xlsx(disgenet, file = "../data/chf_targets/disgenet.xlsx")
```

## GeneCards
```{r}
rm(list = ls())

genecards.hf <- read.csv("../data/chf_targets/genecards/genecards_hf.csv",skip = 4)

genecards.chf <- read.csv("../data/chf_targets/genecards/genecards_chf.csv",skip = 4)

load("../data/uniprot/uniprot_human_reviewed.RData")
uniprot_human_reviewed <- uniprot_human_reviewed |>
  mutate(Gene.Names.primary = trimws(Gene.Names.primary),
         Gene.Names = trimws(Gene.Names)) |>
  dplyr::filter(Gene.Names.primary != "") |>
  dplyr::select(Gene.Names.primary,Gene.Names) |>
  separate_rows(Gene.Names, sep = "\\s+") |>
  dplyr::filter(Gene.Names !="") |>
  mutate(Gene.Names.primary = trimws(Gene.Names.primary),
         Gene.Names = trimws(Gene.Names))

genecards <- rbind(genecards.hf,genecards.chf) |>
  dplyr::group_by(Symbol) |>
  dplyr::slice_max(Relevance.Score, n = 1, with_ties = FALSE) |>
  dplyr::ungroup() |>
  dplyr::filter(!is.na(Symbol)) |>
  mutate(Source = "GeneCards") |>
  dplyr::select(Symbol,Source,Relevance.Score)

nrow(genecards)

save(genecards,file = "../data/chf_targets/genecards.RData")
```

<!-- ## OMIM -->
<!-- ```{r} -->
<!-- rm(list = ls()) -->

<!-- omim <- read.xlsx("../data/chf_targets/omim/omim2.xlsx",startRow = 5) |> -->
<!--   dplyr::filter(!is.na(Approved.Symbol)) |> -->
<!--   distinct(Approved.Symbol,.keep_all = TRUE) |> -->
<!--   dplyr::select(Symbol = Approved.Symbol, Pheno.map.key) |> -->
<!--   distinct(Symbol) |> -->
<!--   dplyr::filter(!is.na(Symbol)) |> -->
<!--   mutate(Source = "OMIM") -->

<!-- nrow(omim) -->

<!-- save(omim,file = "../data/chf_targets/omim.RData") -->

<!-- write.xlsx(omim, file = "../data/chf_targets/omim.xlsx") -->

<!-- omim_mapping <- fread("../data/chf_targets/omim00.tsv.gz") |> -->
<!--   distinct(`Gene Names (primary)`) -->
<!-- ``` -->

## TTD
```{r}
rm(list = ls())

load("../data/uniprot/uniprot_human_reviewed.RData")
uniprot_human_reviewed <- uniprot_human_reviewed |>
  mutate(Gene.Names.primary = trimws(Gene.Names.primary),
         Gene.Names = trimws(Gene.Names)) |>
  dplyr::filter(Gene.Names.primary != "") |>
  dplyr::select(Gene.Names.primary,Gene.Names) |>
  separate_rows(Gene.Names, sep = "\\s+") |>
  dplyr::filter(Gene.Names !="") |>
  mutate(Gene.Names.primary = trimws(Gene.Names.primary),
         Gene.Names = trimws(Gene.Names))

ttd <- read.xlsx("../data/chf_targets/ttd/ttd.xlsx") |>
  dplyr::distinct(Symbol) |>
  left_join(uniprot_human_reviewed, 
            by = c("Symbol" = "Gene.Names"),
            relationship = "many-to-many") |>
  dplyr::filter(!is.na(Gene.Names.primary)) |>
  distinct(Gene.Names.primary) |>
  dplyr::rename(Symbol = Gene.Names.primary) |>
  mutate(Source="TTD")

nrow(ttd)

save(ttd,file = "../data/chf_targets/ttd.RData")
```


## CHF-related targets
```{r}
rm(list = ls())

load("../data/chf_targets/disgenet.RData")
load("../data/chf_targets/genecards.RData")
# load("../data/chf_targets/omim.RData")
load("../data/chf_targets/ttd.RData")

# 0.2
disgenet.filter <- disgenet |>
  dplyr::filter(max_normalized_score >= 0.2) |>
  distinct(Symbol, .keep_all = TRUE ) |> 
  dplyr::select(Symbol,Source)


genecards.filter <- genecards |>
  dplyr::filter(Relevance.Score >= 3) |>
  dplyr::select(-Relevance.Score) |>
  distinct(Symbol, .keep_all = TRUE ) |>
  dplyr::select(Symbol,Source)

chf.targets <- rbind(disgenet.filter, genecards.filter,  ttd) |>
  group_by(Symbol) |>
  summarise(
    Source = paste(Source, collapse = ";"),
    .groups = "drop"
  )

cat("过滤后疾病靶点总数：", nrow(chf.targets))

save(chf.targets, file = "../data/chf_targets/chf.targets.RData")

write.xlsx(chf.targets, file = "supplementary/Table S3.xlsx")
```

# 靶点交集
```{r}
rm(list = ls())

load("../data/chf_targets/chf.targets.RData")
load("../data/qfd_targets/ingredients_targets.RData")

overlapping.targets <- intersect(chf.targets$Symbol, ingredients.targets.filter$Symbol)

overlapping.targets2 <- data.frame(Symbol = overlapping.targets) |>
  left_join(ingredients.targets.filter,by = "Symbol") |>
  distinct(Abbreviation,Symbol)  |>
  group_by(Symbol) |>
  summarise(
    Source = paste(Abbreviation, collapse = ";"),
    .groups = "drop"
  ) |>
  dplyr::select(Abbreviation = Source, Symbol) |>
  arrange(Abbreviation)

cat("治疗靶点总数：", nrow(overlapping.targets2), "\n")
cat("共有靶点数：", sum(overlapping.targets2$Abbreviation == "ALRP;AR"),"\n")
cat("附子特有靶点数：", sum(overlapping.targets2$Abbreviation == "ALRP"), "\n")
cat("黄芪特有靶点数：", sum(overlapping.targets2$Abbreviation == "AR"), "\n")

ingredients.overlapping <- ingredients.targets.filter |>
  dplyr::filter(Symbol %in% overlapping.targets)

length(unique(ingredients.overlapping$PubChem.CID))
length(unique(ingredients.overlapping$Symbol))

ingredients.overlapping.validated <- ingredients.overlapping |>
  dplyr::filter(Class == "validated") 
length(unique(ingredients.overlapping.validated$PubChem.CID))
length(unique(ingredients.overlapping.validated$Symbol))

ingredients.overlapping.predicted <- ingredients.overlapping |>
  dplyr::filter(Class == "predicted") 
length(unique(ingredients.overlapping.predicted$PubChem.CID))
length(unique(ingredients.overlapping.predicted$Symbol))

save(overlapping.targets,file = "../data/network/overlapping.targets.RData")

write.xlsx(data.frame(Symbol=overlapping.targets),file = "../data/network/overlapping.targets.xlsx")

write.xlsx(data.frame(Symbol = overlapping.targets),file = "supplementary/Table S4.xlsx")


# 1. 提取三组基因符号
alrp_unique <- overlapping.targets2$Symbol[overlapping.targets2$Abbreviation == "ALRP"]
ar_unique <- overlapping.targets2$Symbol[overlapping.targets2$Abbreviation == "AR"]
common <- overlapping.targets2$Symbol[overlapping.targets2$Abbreviation == "ALRP;AR"]

# 2. 查看各组数量，确认无误
cat("附子独有：", length(alrp_unique), "\n")
cat("黄芪独有：", length(ar_unique), "\n")
cat("共有：", length(common), "\n")
cat("总计：", length(alrp_unique) + length(ar_unique) + length(common), "\n")

# 3. 保存为RData，方便后续分析
save(alrp_unique, ar_unique, common, overlapping.targets2, 
     file = "../data/network/final_261_targets_groups.RData")
```


# 参考文献 {-}
